---
title: 'Class 6a: Blake et al. (2015) exercise'
author: "Mike Frank"
date: "November 16, 2017"
output: 
  html_document:
    toc: true
---
  
# Intro
  
This is an exploration of Blake et al. (2015), [Ontogeny of fairness in seven societies](http://www.nature.com/nature/journal/v528/n7581/full/nature15703.html).

```{r}
library(tidyverse)

d <- read_csv("data/Ontogeny_fairness_seven_societies_data.csv") %>%
  mutate(trial_num = as.numeric(gsub(".(\\d+)", "\\1", trial)),
         trial_type = factor(eq.uneq, 
                             levels = c("E","U"), 
                             labels = c("Equal","Unequal")), 
         condition = factor(condition,
                            levels = c("AI","DI"), 
                            labels = c("Advantageous","Disadvantageous"))) %>%
  arrange(country, condition, actor.id, trial_num)
```

The general structue of this experiment is that we have pairs of participants assigned to the Disadvantageous and Advantageous conditions, and they do 16 trials for each pair in two blocks (varying value of the offer). 

```{r}
d
```


# Variable exploration

Describe the dataset graphically in ways that are useful for you to get a handle on the data collection effort. 

```{r}

```

# Hypothesis-related exploration

Try plotting country data so you can explore!

```{r}

```

Now apply this approach to all countries using faceting.

```{r}

```

