---
title: "Reproducibility check: [Author et al. (Year)]"
author: "[Your name]"
date: today
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Article

- Citation and DOI:
- Data/materials repository URL:
- Part A (known case) or Part B (class corpus):

## Target outcomes

Written **before** opening the data. Quote each from the paper.

| # | Target outcome (quoted) | Paper value(s) |
|---|---|---|
| 1 | | |
| 2 | | |
| 3 | | |

## Materials inventory

| Item | Available? | Notes |
|---|---|---|
| Raw data | | |
| Processed data | | |
| Codebook / variable descriptions | | |
| Analysis code | | |
| Preregistration | | |

## Reproduction

The agent's work, under your direction. Keep all code visible.

```{r}
#| label: setup
library(tidyverse)
```

```{r}
#| label: load
```

```{r}
#| label: preprocess
```

```{r}
#| label: analysis
```

## Checker's verdict

Filled in by you, not the agent.

| # | Paper value | Reproduced value | Verdict (exact / minor / major / not reproducible) | Reason for discrepancy |
|---|---|---|---|---|
| 1 | | | | |
| 2 | | | | |
| 3 | | | | |

Agent's own claim about success, and whether it was right:

## Time

- Your time (hours):
- Agent wall-clock time (hours):
- Stopped at the 4-hour cap? What remained:

## Assessment

Is the primary result analytically reproducible? What would have made this take 30 minutes?

## AI documentation

- Agent and interface (e.g., Claude Code, Codex CLI, Cursor):
- Model and version, if the tool reports it:
- Context you gave it (files, instructions, any context file):
- Number of times you intervened, and the most important intervention:
- Transcript: link to the exported transcript in your repo (`transcripts/`). If your tool
  cannot export one, paste your prompts in order here instead.
