---
title: "Reproducibility Report: Group B Choice 3"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

-------

For this exercise, please try to reproduce the results from Experiment 2 of the associated paper (de la Fuente, Santiago, Roman, Dumitrache, & Casasanto, 2014). The PDF of the paper is included in the same folder as this Rmd file. 

#### Methods summary: 

Researchers tested the question of whether temporal focus differs between Moroccan and Spanish cultures, hypothesizing that Moroccans are more past-focused, whereas Spaniards are more future-focused. Two groups of participants ($N = 40$ Moroccan and $N=40$ Spanish) completed a temporal-focus questionnaire that contained questions about past-focused ("PAST") and future-focused ("FUTURE") topics. In response to each question, participants provided a rating on a 5-point Likert scale on which lower scores indicated less agreement and higher scores indicated greater agreement. The authors then performed a mixed-design ANOVA with agreement score as the dependent variable, group (Moroccan or Spanish, between-subjects) as the fixed-effects factor, and temporal focus (past or future, within-subjects) as the random effects factor. In addition, the authors performed unpaired two-sample t-tests to determine whether there was a significant difference between the two groups in agreement scores for PAST questions, and whether there was a significant difference in scores for FUTURE questions.

------

#### Target outcomes: 

Below is the specific result you will attempt to reproduce (quoted directly from the results section of Experiment 2):

> According to a mixed analysis of variance (ANOVA) with group (Spanish vs. Moroccan) as a between-subjects factor and temporal focus (past vs. future) as a within-subjectS factor, temporal focus differed significantly between Spaniards and Moroccans, as indicated by a significant interaction of temporal focus and group, F(1, 78) = 19.12, p = .001, ηp2 = .20 (Fig. 2). Moroccans showed greater agreement with past-focused statements than Spaniards did, t(78) = 4.04, p = .001, and Spaniards showed greater agreement with future-focused statements than Moroccans did, t(78) = −3.32, p = .001. (de la Fuente et al., 2014, p. 1685).

------

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

# Step 1: Load packages

```{r}
library(tidyverse) # for data munging
library(knitr) # for kable table formating
library(haven) # import and export 'SPSS', 'Stata' and 'SAS' Files
library(readxl) # import excel files
library(afex)
library(heplots)

# #optional packages/functions:
 library(afex) # anova functions
# library(ez) # anova functions 2
 library(scales) # for plotting
# std.err <- function(x) sd(x)/sqrt(length(x)) # standard error
```

# Step 2: Load data

```{r}
# Just Experiment 2
data_path <- 'data/DeLaFuenteEtAl_2014_RawData.xls'
d <- read_excel(data_path, sheet=3)
```

# Step 3: Tidy data

```{r}
#data looks tidy already - no changes attempted
```

# Step 4: Run analysis

## Pre-processing

```{r}
d2 <- d %>%
  mutate(id=ifelse(group=="Moroccan",participant,participant+40)) %>%
  arrange(group, participant,item) %>%
  rename(rating= "Agreement (0=complete disagreement; 5=complete agreement)") %>%
  distinct(id, group, participant, subscale, item, .keep_all=T)

d2$group <- str_replace_all(d2$group, "young Spaniard", "Spaniard")
```

## Descriptive statistics

Try to recreate Figure 2 (fig2.png, also included in the same folder as this Rmd file):

```{r}
d2_graph = d2 %>%
  group_by(group, subscale) %>%
  summarise(mean_rating = mean(rating, na.rm=T),
            n=n(),
            se_rating = sd(rating)/sqrt(n))

d2_graph$group <- factor(d2_graph$group,levels=c("Spaniard", "Moroccan"))

ggplot(d2_graph, aes(x=group, y=mean_rating, fill=subscale)) +
  geom_bar(position="dodge", stat="identity") +
  scale_y_continuous(limits=c(2.00, 4.00),oob = rescale_none) +
  geom_errorbar(aes(ymin=mean_rating-se_rating, ymax=mean_rating+se_rating), width = .5, position=position_dodge(.9), na.rm=T) +
  labs(y="Rating", x = "Group") 
```
## Inferential statistics

> According to a mixed analysis of variance (ANOVA) with group (Spanish vs. Moroccan) as a between-subjects factor and temporal focus (past vs. future) as a within-subjects factor, temporal focus differed significantly between Spaniards and Moroccans, as indicated by a significant interaction of temporal focus and group, F(1, 78) = 19.12, p = .001, ηp2 = .20 (Fig. 2).

```{r}
# reproduce the above results here

d3 <- d2 %>%
  group_by(id, group, participant, subscale) %>%
  summarise(rating=mean(rating))  #mean rating by participant for past/future subscale

d3$id <- factor(d3$id)
Within.aov.1 <- aov_car(rating ~ group*subscale + Error(id/subscale), data=d3, anova_table=list(es="pes"), na.rm = TRUE)
summary(Within.aov.1)

model.d <- aov(rating ~ as.factor(subscale)*as.factor(group), data=d3)
summary(model.d)

etasq(model.d, anova = FALSE, partial = TRUE)

```

> Moroccans showed greater agreement with past-focused statements than Spaniards did, t(78) = 4.04, p = .001, 

```{r}
# reproduce the above results here
d4 <- d3 %>%
  filter(subscale=="FUTURE")
t.test(rating ~ group, data = d4)
```

> and Spaniards showed greater agreement with future-focused statements than Moroccans did, t(78) = −3.32, p = .001.(de la Fuente et al., 2014, p. 1685)

```{r}
# reproduce the above results here
d5 <- d3 %>%
  filter(subscale=="PAST")
t.test(rating ~ group, data = d5)
```

# Step 5: Reflection

Were you able to reproduce the results you attempted to reproduce? If not, what part(s) were you unable to reproduce?

> I was only able to reproduce the figure 2. I was unable to reproduce other findings.

How difficult was it to reproduce your results? 

> It was very difficult to reproduce the results. I was unsuccessful in doing so. I also surpassed the 3 hour mark, so I decided to stop working on the assignmnet as directed by Mike.

What aspects made it difficult? What aspects made it easy?

> It's hard to tell what specifically made it difficult. I wasn't able to work with other members of the class on this assignment as much as I was able to do with my Group A choice 3 assignment. This really affected my abilitiy to progress. For someone who really does not have a strong statistics or R background, completing an assignmnet like this on my own is nearly impossible (especially in a 3 hour block of time)
